### 2020-07-30 Model Data Sheets and Code/Manuscript_Code.Rmd
New document with final code for manuscript

---
title: "Zaca Fire Type Conversion"
date: "7/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Prepare Workspace

```{r load_libraries}

library(tidyverse)
library(raster)
library(VSURF)
library(ALEplot)

setwd("workingdirectory") #change this to your local directory

set.seed(2020) #Set seed so random forests are replicable

```

### Load Shapefiles

```{r load_shapefiles}

zaca <- shapefile("E:/zaca/chris/zaca_scar")

mixed_conifer <- shapefile("all_mixed_pine.shp")

coulter <- shapefile("coulter_pines.shp") %>% 
  spTransform(., crs(mixed_conifer))

```

### Prepare Environmental Variables

```{r}

dem <- raster("E:/zaca/environmental_data/all_variables/dem_10m_zaca.tif")

vpd_jan <- raster("E:/zaca/environmental_data/cwd_vpd_amp/PRISM_vpdmax_30yr_normal_800mM2_01_asc/PRISM_vpdmax_30yr_normal_800mM2_01_asc.asc") %>% 
  projectRaster(., dem_30m) %>% 
  crop(., zaca) %>% 
  mask(., zaca)

#TPI
#Insolation
#Temp
#Rainfall

```

## Mixed Conifer Forest

### Extract Values

```{r}

mixed_conifer_env <- extract()

```

### Stepwise Variable Selection

```{r}

#Extract prediction and response variables
mcf.predictors <- mixed_conifer_env[,2:ncol(mixed_conifer_env)]
mcf.response <- mixed_conifer_env[,1]

#Run stepwise variable selection algorithm
mcf.vsurf <- VSURF(x = predictors, y = response)

#Get names of selected variables
names(mcf.predictors[mcf.vsurf$varselect.pred])

```

### Random Forest

```{r}

#Create random forest using selected variables
mixed_conifer_rf <- randomForest(RdNBR_2007 ~ var1 + var2 + var3, 
                                 data = mixed_conifer_env, 
                                 importance = TRUE, 
                                 na.action = na.omit)

```