### 2020-08-02 Model Data Sheets and Code/Manuscript_Code.Rmd
Calculate fire history

---
title: "Zaca Fire Type Conversion"
date: "7/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prepare Workspace

```{r load_libraries}

library(tidyverse)
library(raster)
library(randomForest)
library(VSURF)
library(ALEplot)

setwd("E:/zaca/type_conversion")

set.seed(2020) #Set seed so random forests are replicable

UTM_11N <- CRS("+proj=utm +zone=11 +ellps=WGS84 +datum=WGS84 +units=m +no_defs ")

```

## Prepare Vector Data

### Fire History

#### Load Fire History Data

```{r}

#Load Fires
firep19_1 <- shapefile("./fire19_1/firep19_1.shp")

#Rename Columns
names(firep19_1@data)[which(names(firep19_1@data) == "ALARM_DATE")] <- "START_DATE"
names(firep19_1@data)[which(names(firep19_1@data) == "CONT_DATE")] <- "END_DATE"
names(firep19_1@data)[which(names(firep19_1@data) == "YEAR_")] <- "YEAR"

#Load Prescribed Burns
rxburn19_1 <- shapefile("./fire19_1/rxburn19_1.shp")

#Rename Columns
names(rxburn19_1@data)[which(names(rxburn19_1@data) == "TREATMENT1")] <- "FIRE_NAME"
names(rxburn19_1@data)[which(names(rxburn19_1@data) == "YEAR_")] <- "YEAR"

#Combine Layers
fire_history <- bind(firep19_1, rxburn19_1)

#Change Column Data Types
fire_history$YEAR <- as.numeric(fire_history$YEAR)

#Select Columns
fire_history <- fire_history[, c("FIRE_NAME", "YEAR")]

```

#### Calculate Years Since Fire

```{r}

#Extract Zaca burn scar
zaca <- subset(fire_history, FIRE_NAME == "ZACA" & YEAR == 2007)

#Extract fires before Zaca
fires_before_zaca <- subset(fire_history, YEAR < 2007)

#Calculate years since last burn
fires_within_zaca_scar <- intersect(zaca, fires_before_zaca)
fires_within_zaca_scar$years_since_fire <- fires_within_zaca_scar$YEAR.1 - fires_within_zaca_scar$YEAR.2

#Reproject to UTM Zone 11N
fires_within_zaca_scar <- spTransform(fires_within_zaca_scar, UTM_11N)

```

#### Reproject Zaca Boundary

```{r}

zaca <- spTransform(zaca, UTM_11N)

```

### Vegetation Cover

```{r}

south_coast_vegetation <- shapefile("RVMid_R05_SouCoast.shp") %>% 
  crop(., zaca)

central_coast_vegetation <- shapefile("RVMid_R05_CentralCoast.shp") %>% 
  crop(., zaca)

zaca_vegetation <- merge(south_coast_vegetation, central_coast_vegetation)

```

## Prepare Environmental Variables

## Create Raster Templates

The templates have the correct projection, extent, origin, and spatial resolution for the environmental variables. The environmental variables will be reprojected to line up with the templates. The pixel grid is derived from the Landsat images. The topographic variables and insolation will be analyzed at a 10 m spatial resolution. The other variables will be analyzed at a 30 m spatial resolution.

```{r}

template_30m <- raster(landsat_2006[[1]]) %>% 
  crop(., y = zaca)

template_10m <- raster(landsat_2006[[1]]) %>% 
  crop(., y = zaca) %>% 
  disaggregate(., fact = 3)

```

### Prepare Digital Elevation Model

```{r dem}

dem120 <- raster("C:/Users/Christopher/Downloads/USGS_13_n35w120.tif")
dem121 <- raster("C:/Users/Christopher/Downloads/USGS_13_n35w121.tif")

usgs_dem <- merge(dem120, dem121) %>% 
  projectRaster(., to = template_10m, method = "bilinear")

```

### Prepare Topography Data

```{r topography}

```

### Prepare Climate Data

```{r climate}

dem <- raster("E:/zaca/environmental_data/all_variables/dem_10m_zaca.tif")

vpd_jan <- raster("E:/zaca/environmental_data/cwd_vpd_amp/PRISM_vpdmax_30yr_normal_800mM2_01_asc/PRISM_vpdmax_30yr_normal_800mM2_01_asc.asc") %>% 
  projectRaster(., dem_30m) %>% 
  crop(., zaca) %>% 
  mask(., zaca)

#TPI
#Insolation
#Temp
#Rainfall

```

## Mixed Conifer Forest

### Extract Values

```{r}

mixed_conifer_env <- extract()

```

### Stepwise Variable Selection

```{r}

#Extract prediction and response variables
mcf.predictors <- mixed_conifer_env[,2:ncol(mixed_conifer_env)]
mcf.response <- mixed_conifer_env[,1]

#Run stepwise variable selection algorithm
mcf.vsurf <- VSURF(x = predictors, y = response)

#Get names of selected variables
names(mcf.predictors[mcf.vsurf$varselect.pred])

```

### Random Forest

```{r}

#Create random forest using selected variables
mixed_conifer_rf <- randomForest(RdNBR_2007 ~ var1 + var2 + var3, 
                                 data = mixed_conifer_env, 
                                 importance = TRUE, 
                                 na.action = na.omit)

```