### 2020-08-08 chi_sq for vegetation and seedling age classes.Rmd
analysis for field sruvey results section

---
title: "Field Surveys"
author: "Anne-Marie Parkinson"
date: "August 7, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
# create data frame
df <- data.frame(age_class = c("0-3", "4-6", "7-10"),
                 cepa = c(1, 6, 41),
                 chaparral = c(12, 13, 27))

# make contigency table: getting the data in a format that will be used in chi.sq()
df_contigency <- df %>% 
  column_to_rownames('age_class')

# run chi-square test
df_chi <- chisq.test(df_contigency)
df_chi # pvalue = 0.000665

# get Standardized residuals
df_chi$stdres
```
The numbers reported in the above data frame come from the manuscript, which I calcualted from excel sheets.
cepa = the number of recorded sapling/seedlings found in ceanothus palmeri in each corresponding age class 
chaparral = the number of recorded sapling/seedlings found in mixed chaparral in each corresponding age class

Standardized residuals > |2| indicates strong divergence from null hypothesis scenario. So There are significantly more seedlings/saplings in the 7-10 age class that were found in ceanothus palmeri, But both vegetation types had a majority of seedling/saplings in the 7-10 age class. And there were significantly more seedlings/saplings in the 0-3 age class that were found in mixed chaparral.

($\chi$^2^(`r df_chi$parameter`) = `r round(df_chi$statistic,2)`, *p* = `r round(df_chi$p.value,2)`)

chris, later we can combine this with the final manuscript code.