### 2020-10-09 Model Data Sheets and Code/Manuscript_Code.Rmd
work on combined model

---
title: "Zaca Fire Type Conversion"
date: "7/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prepare Workspace

```{r load_libraries}

library(tidyverse)
library(raster)
library(randomForest)
library(VSURF)
library(ALEPlot)
library(tuneRanger)
library(iml)
library(ggpubr)

setwd("E:/zaca/type_conversion")

set.seed(2020) #Set seed so random forests are replicable

UTM_11N <- CRS("+proj=utm +zone=11 +ellps=WGS84 +datum=WGS84 +units=m +no_defs ")

```

## Prepare Imagery Data

### Load Function to Pre-process Landsat

The Landsat pre-processing function is stored in a separate file. For each image, it stacks the Landsat bands into a single image file. It then masks erroneous reflectance retrievals, masks clouds, and saves the new image. The function automatically iterates over every data set in a subdirectory.

```{r}

source("preprocessLandsat.R")

```

### Pre-process Landsat Images

```{r}

preprocessLandsat("E:/zaca/type_conversion/Imagery")

```

### Calculate NBR

```{r}

#Directory will all pre-processed Landsat images
setwd("E:/zaca/type_conversion/Imagery") #changed from: setwd("E:/zaca/type_conversion/Imagery/Level2")

for(img in list.files(path = ".", "tif$")){ # changed from for(img in list.files(path = ".", "GTiff$"))

  print(paste("Processing ", substr(img, 18, 25), sep = ""))

  #Merge bands into a brick
  image <- brick(img)

  NAvalue(image) <- -9999

  #Check sensor ID, mask using QA layer, and then compute NBR using appropriate formula
  if(as.numeric(substr(img, 3, 4)) == 8){

    print("Recognized Landsat 8")

    #Band designations change, depending on whether the OLI aerosol band is included
    if(nlayers(image) == 7){nbr <- (image[[5]] - image[[7]])/(image[[5]] + image[[7]])}
    else if(nlayers(image) == 6){nbr <- (image[[4]] - image[[6]])/(image[[4]] + image[[6]])}
    else {print("Incorrect number of bands")}

  } else if(as.numeric(substr(img, 3, 4)) == 5 | as.numeric(substr(img, 3, 4)) == 7) {

    print("Recognized Landsat 5 or 7")

    #Band 7 is the sixth band in the brick for TM/ETM
    if(nlayers(image) == 6){nbr <- (image[[4]] - image[[6]])/(image[[4]] + image[[6]])}
    else {print("Incorrect number of bands")}

    } else {print("Error recognizing sensor")}

  writeRaster(nbr, filename = paste("NBR_", substr(img, 18, 25), ".tif", sep = ""), format = "GTiff", NAflag = -9999)
  assign(paste("NBR_", substr(img, 18, 25)), nbr)

  rm(image, image_brick, nbr)

  print("Image completed")

}

```

### Calculate RdNBR

```{r}

NBR_brick <- brick(NBR_2000, NBR_2001, NBR_2002, NBR_2003, NBR_2004, NBR_2005, NBR_2006, NBR_2007) * 1000

baseline_NBR <- calc(NBR_brick[[1:7]], fun = median)

RdNBR_2007 <- (baseline_NBR - NBR_brick[[8]]) / sqrt(abs(baseline_NBR/1000))

```

## Prepare Vector Data

### Fire History

#### Load Fire History Data

```{r}

#Load Fires
firep19_1 <- shapefile("./fire19_1/firep19_1.shp")

#Rename Columns
names(firep19_1@data)[which(names(firep19_1@data) == "ALARM_DATE")] <- "START_DATE"
names(firep19_1@data)[which(names(firep19_1@data) == "CONT_DATE")] <- "END_DATE"
names(firep19_1@data)[which(names(firep19_1@data) == "YEAR_")] <- "YEAR"

#Load Prescribed Burns
rxburn19_1 <- shapefile("./fire19_1/rxburn19_1.shp")

#Rename Columns
names(rxburn19_1@data)[which(names(rxburn19_1@data) == "TREATMENT1")] <- "FIRE_NAME"
names(rxburn19_1@data)[which(names(rxburn19_1@data) == "YEAR_")] <- "YEAR"

#Combine Layers
fire_history <- bind(firep19_1, rxburn19_1)

#Change Column Data Types
fire_history$YEAR <- as.numeric(fire_history$YEAR)

#Select Columns
fire_history <- fire_history[, c("FIRE_NAME", "YEAR")]

```

#### Calculate Years Since Fire

```{r}

#Extract Zaca burn scar
zaca <- subset(fire_history, FIRE_NAME == "ZACA" & YEAR == 2007)

#Extract fires before Zaca
fires_before_zaca <- subset(fire_history, YEAR < 2007)

#Calculate years since last burn
fires_within_zaca_scar <- intersect(zaca, fires_before_zaca)
fires_within_zaca_scar$years_since_fire <- fires_within_zaca_scar$YEAR.1 - fires_within_zaca_scar$YEAR.2

#Reproject to UTM Zone 11N
fires_within_zaca_scar <- spTransform(fires_within_zaca_scar, UTM_11N)

```

#### Reproject Zaca Boundary

```{r}

zaca <- spTransform(zaca, UTM_11N)

```

### Vegetation Cover

```{r}

south_coast_vegetation <- shapefile("RVMid_R05_SouCoast.shp") %>% 
  crop(., zaca)

central_coast_vegetation <- shapefile("RVMid_R05_CentralCoast.shp") %>% 
  crop(., zaca)

zaca_vegetation <- merge(south_coast_vegetation, central_coast_vegetation)

```

## Prepare Environmental Variables

## Create Raster Templates

The templates have the correct projection, extent, origin, and spatial resolution for the environmental variables. The environmental variables will be reprojected to line up with the templates. The pixel grid is derived from the Landsat images. The topographic variables and insolation will be analyzed at a 10 m spatial resolution. The other variables will be analyzed at a 30 m spatial resolution.

```{r}

template_30m <- raster(landsat_2006[[1]]) %>% 
  crop(., y = zaca)

template_10m <- raster(landsat_2006[[1]]) %>% 
  crop(., y = zaca) %>% 
  disaggregate(., fact = 3)

```

### Prepare Digital Elevation Model

```{r dem}

dem120 <- raster("C:/Users/Christopher/Downloads/USGS_13_n35w120.tif")
dem121 <- raster("C:/Users/Christopher/Downloads/USGS_13_n35w121.tif")

usgs_dem <- merge(dem120, dem121) %>% 
  projectRaster(., to = template_10m, method = "bilinear")

#dem <- raster("E:/zaca/environmental_data/all_variables/dem_10m_zaca.tif")

```

### Prepare Topography Data

```{r topography}

slope <- terrain(x = usgs_dem, opt = "slope", unit = "degrees", neighbors = 8)

aspect <- terrain(x = usgs_dem, opt = "aspect", unit = "degrees", neighbors = 8)

tpi <- terrain(x = usgs_dem, opt = "TPI") #Can calculate larger moving windows manually

```

### Prepare Climate Data

```{r climate}

jan_max_vpd_30yr <- raster("E:/zaca/environmental_data/cwd_vpd_amp/PRISM_vpdmax_30yr_normal_800mM2_01_asc/PRISM_vpdmax_30yr_normal_800mM2_01_asc.asc") %>% 
  projectRaster(., template_30m, method = "ngb") %>% 
  crop(., zaca) %>% 
  mask(., zaca)

jan_min_temp_30yr <- raster("E:/zaca/type_conversion/Environmental_Variables/Climate/January_Min_Temperature/PRISM_tmin_30yr_normal_800mM2_01_asc/PRISM_tmin_30yr_normal_800mM2_01_asc.asc") %>% 
  projectRaster(., template_30m, method = "ngb") %>% 
  crop(., zaca) %>% 
  mask(., zaca)

aug_max_temp_30yr <- raster("E:/zaca/type_conversion/Environmental_Variables/Climate/August_Max_Temperature/PRISM_tmax_30yr_normal_800mM2_08_asc/PRISM_tmax_30yr_normal_800mM2_08_asc.asc") %>% 
  projectRaster(., template_30m, method = "ngb") %>% 
  crop(., zaca) %>% 
  mask(., zaca)

aug_max_vpd_30yr <- raster("E:/zaca/type_conversion/Environmental_Variables/Climate/August_Max_VPD/PRISM_vpdmax_30yr_normal_800mM2_08_asc/PRISM_vpdmax_30yr_normal_800mM2_08_asc.asc") %>% 
  projectRaster(., template_30m, method = "ngb") %>% 
  crop(., zaca) %>% 
  mask(., zaca)

annual_precip_30yr <- raster("E:/zaca/type_conversion/Environmental_Variables/Climate/Annual_Precipitation/PRISM_ppt_30yr_normal_800mM2_annual_asc/PRISM_ppt_30yr_normal_800mM2_annual_asc.asc") %>% 
  projectRaster(., template_30m, method = "ngb") %>% 
  crop(., zaca) %>% 
  mask(., zaca)

#Add Insolation

```

### Prepare Weather Data

```{r}

jan_min_temp_2007 <- raster("E:/zaca/type_conversion/Environmental_Variables/Weather/January_Min_Temp/PRISM_tmin_stable_4kmM3_200701_bil/PRISM_tmin_stable_4kmM3_200701_bil.bil") %>% 
  projectRaster(., template_30m, method = "ngb") %>% 
  crop(., zaca) %>% 
  mask(., zaca)

aug_max_temp_2006 <- raster("E:/zaca/type_conversion/Environmental_Variables/Climate/August_Max_Temperature/PRISM_tmax_30yr_normal_800mM2_08_asc/PRISM_tmax_30yr_normal_800mM2_08_asc.asc") %>% 
  projectRaster(., template_30m, method = "ngb") %>% 
  crop(., zaca) %>% 
  mask(., zaca)

aug_max_vpd_2006 <- raster("E:/zaca/type_conversion/Environmental_Variables/Weather/August_Max_VPD/PRISM_vpdmax_stable_4kmM3_200608_bil/PRISM_vpdmax_stable_4kmM3_200608_bil.bil") %>% 
  projectRaster(., template_30m, method = "ngb") %>% 
  crop(., zaca) %>% 
  mask(., zaca)

annual_precip_30yr <- raster("E:/zaca/type_conversion/Environmental_Variables/Climate/Annual_Precipitation/PRISM_ppt_30yr_normal_800mM2_annual_asc/PRISM_ppt_30yr_normal_800mM2_annual_asc.asc") %>% 
  projectRaster(., template_30m, method = "ngb") %>% 
  crop(., zaca) %>% 
  mask(., zaca)

#Calculate precipitation for preceding year?

```

## Mixed Conifer Forest

###... Load data

the above code wouldnt work, so I got evi and min jan temp in the aca mcf mortality model results.rmd and loaded the old files ive been using
```{r}
mcf <- read.csv(here::here("Model Data Sheets and Code", 'Zaca MCF variables for model.csv')) %>% 
  mutate(mortality_qual= fct_relevel(mortality_qual, levels=c("High", "Moderate", "Low", "None"))) %>% 
  mutate(aspect_transformed = (1-cos((2*pi*aspect)/360))/2) %>% 
  mutate(two_class_mortality = case_when(
    mortality_qual %in% "High" ~ "High",
    mortality_qual %in% c("Moderate", "Low", "None") ~ "NLM")) %>% 
  mutate(two_class_mortality = as.factor(two_class_mortality),
         two_class_mortality= fct_relevel(two_class_mortality, levels=c("NLM", "High"))) %>% 
  mutate(tc_mortality = as.factor(tc_mortality)) %>% 
  mutate(rdnbr_qualitative = case_when(
    rdnbr_2007 < 640 ~ "NLM",
    rdnbr_2007 < 850 ~ "High",
    rdnbr_2007 >= 850 ~ "Type Converted")) %>%
  mutate(rdnbr_qualitative = as.factor(rdnbr_qualitative),
         rdnbr_qualitative = fct_relevel(rdnbr_qualitative, levels=c("NLM","High", "Type Converted"))) %>% 
  #   mutate(rdnbr_qualitative2 = case_when(
  #   rdnbr_qualitative %in% "High" ~ "High",
  #   rdnbr_qualitative %in% c("Moderate", "Low", "None") ~ "NLM")) %>% 
  # mutate(rdnbr_qualitative2 = as.factor(rdnbr_qualitative2),
  #        rdnbr_qualitative2= fct_relevel(rdnbr_qualitative2, levels=c("NLM", "High")))%>% 
  mutate(rdnbr_tc = case_when(
    rdnbr_2007 < 850 ~ "Not_TC",
    rdnbr_2007 >= 850 ~ "TC"
  )) %>% 
  mutate(rdnbr_tc = as.factor(rdnbr_tc)) %>% 
  mutate(rdnbr_tc= fct_relevel(rdnbr_tc, levels=c("Not_TC", "TC")))

```

### ... summary statistics
```{r}
# sample sizes
mcf_tc_samplesize <- mcf %>% group_by(tc_mortality) %>%count() #type converted manual estimates n = 99
mcf_tc_samplesie <- mcf %>% group_by(rdnbr_tc) %>%dplyr::count() # rdnbr=1000 for type converted  n = 55 (almost too small to use)
mcf_tc_samplesie <- mcf %>% group_by(rdnbr_qualitative) %>%dplyr::count() #rdnbr = 850:  high=66, low= 585, moderate=272, type converted n = 84

# acres by burn severity class an dtotal acres
mcf_tc_samplesie <- mcf %>% group_by(rdnbr_qualitative) %>%dplyr::summarise(acres = sum(Shape_Area/4047)) %>%dplyr::summarise(tot_acres = sum(acres))

# look at correlations/trends in potentially correlated variables
ggplot(mcf, aes(x = rdnbr_tc, y = rdnbr_2007)) +geom_jitter(aes(color=tc_mortality))
ggplot(mcf, aes(x=dnbr_2007, y=rdnbr_2007)) +geom_point(aes(color=tc_mortality)) + geom_hline(yintercept=875)
ggplot(mcf, aes(ndmi_2006, y=evi_2006)) +geom_point()
ggplot(mcf, aes(dem_10m_zaca, y=tmin_jan2007)) +geom_point()
mod <- lm(ndmi_2006~evi_2006, data=mcf) # R2=0.18
mod1 <- lm(dem_10m_zaca~tmin_jan2007, data=mcf) # R2=0.04
mod2 <- lm(dem_10m_zaca~cwd, data=mcf) # R2=0.47
mod3 <- lm(dem_10m_zaca~rain, data=mcf) # R2=0.498
mod4 <- lm(dem_10m_zaca~tmax, data=mcf) # R2=0.5
mod5 <- lm(dem_10m_zaca~vpd_aug, data=mcf) # R2=0.56
mod6 <- lm(tmax~vpd_aug, data=mcf) # R2=0.95
mod7 <- lm(cwd~vpd_aug, data=mcf) # R2=0.53
mod8 <- lm(tmax~cwd, data=mcf) # R2=0.49

# summary info on TSLF adn elevation for the manuscript
mcf %>% dplyr::summarise(min = min(tslf),
                         mean = mean(tslf),
                         median = median(tslf),
                         max = max(tslf), 
                         min_elev = min(dem_10m_zaca),
                         mean_elev = mean(dem_10m_zaca),
                         max_elev = max(dem_10m_zaca))

```
surprisingly, far fewer variables are strongly correlated with each other and elevation than I would've have thought.

###... Extract Values

```{r}

mixed_conifer_env <- extract()

```

###... Stepwise Variable Selection (VSURF)

```{r}

#Extract prediction and response variables
mcf.predictors <- mcf_model[,2:ncol(mcf_model)]
mcf.response <- mcf_model[,1]

#Run stepwise variable selection algorithm
mcf.vsurf <- VSURF(x = mcf.predictors, y = mcf.response)

#Get names of selected variables
names(mcf.predictors[mcf.vsurf$varselect.thres])
names(mcf.predictors[mcf.vsurf$varselect.interp])
names(mcf.predictors[mcf.vsurf$varselect.pred])

```
run 1 results: 
prediction step, variables selected:all

interp step: variables se;ected: "vpd_aug"         "slope"           "rain"            "tmax"            "tmin_jan"        "sr" "aug_tmax2006"    "tmin_jan2007"    "precip_2006"     "aug_max_vpd2006" "ndmi_2006"       "evi_2006"  "tslf"

prediction step, variables selected: "vpd_aug"      "slope"        "rain"         "tmax"         "tmin_jan"     "sr"           "aug_tmax2006"  "tmin_jan2007" "ndmi_2006"    "evi_2006"     "tslf"

so climate variables are performing better than the weather variables (maybe bc the climate data is avail in higher resolution so there is more variability in values?)

*removed aug_max_vpd2006 and precip_2006 and reran VSURF*

run 2 results: 
interp step: "vpd_aug"      "slope"        "rain"         "tmax"         "tmin_jan"     "tmin_jan2007"

predicition step: "vpd_aug"      "slope"        "rain"         "tmax"         "tmin_jan2007"

###... Tuning Parameters

source: https://arxiv.org/pdf/1804.03515.pdf (see digital page 9 for tuneRanger)

options for the measure function: https://mlr.mlr-org.com/articles/tutorial/measures.html

tunRanger can also be used to tune for ntree, but studies have shown that after a certain number of trees, usually a couple hundred, there is no benefit on accuracy/model fit of increasing ntree, so i fgiure just using a large number will get the same result as whatever number tuneRangere spits out. 
```{r}
set.seed(1127)

# select variables to be used in the model 
mcf_model3 <- mcf %>% 
  dplyr::select(rdnbr_qualitative, tmin_jan2007, vpd_aug, cwd, tmax, rain, sr, slope, aspect_transformed, TWI, ndmi_2006, evi_2006, tslf) # not sure if fri and years so fire, whic sounds like Time Since Last Fire, are the same thing

# run tuneRanger
mcf_task <- makeClassifTask(data = mcf_model3, target = "rdnbr_qualitative")
tune_output <- tuneRanger(mcf_task, 
                          measure = list(mmce), 
                          num.trees = 1500, 
                          tune.parameters = c("mtry", "min.node.size"))

# recommended parameter settings
# get best parameter values
tune_output$model

```
get error when I try to add "replace" as a tuning parameter :(

Trials: (2 classes)
- rdnbr cut off: 850; variables included: tmin_jan2007, tmax, vpd_aug, cwd, rain, sr, slope, aspect_transformed, TPI_five, TWI, ndmi_2006, evi_2006, fri; optimal parameters: mtry=6, min.node.size=4; acc = 82%, not tc = 75%, tc=62%, auc=84%; stable VI: kinda
Using this one- rdnbr cut off: 850; variables included: removed tpi, mtry=3,min.node.size=7
- rdnbr cut off: 850; variables included: removed tmin; optimal parameters: mtry=5, min.node.size=10; stable VI: kinda
- rdnbr cut off: 850; tmax, vpd_aug, rain, sr, slope, aspect_transformed, evi_2006, tslf, tmin_jan2007, ndmi_2006; mtry=3,min.node.size=7; stable vi: sure!
- rdnbr cut off: 850; variables included: removed cwd; optimal parameters: mtry=4, min.node.size=2; stable VI: NO! horrible stability. 
- rdnbr cut off: 850; variables included: removed evi_2006; optimal parameters: mtry=4, min.node.size=3; stable VI:kinda
- rdnbr cut off: 850; variables included: removed ndmi_2006; optimal parameters: mtry=6, min.node.size=27; stable VI:kinda

- rdnbr cut off: 875; variables included: all; optimal parameters: mtry=6, min.node.size=15; stable VI:eh, kinda
- rdnbr cut off: 875; variables inlcuded: rdnbr_tc, tmax, vpd_aug, rain, tmin_jan, sr, slope, aspect_transformed, TPI_five, TWI, evi_2006, tslf, cwd, tmin_jan2007, ndmi_2006, precip_2006, aug_max_vpd2006, aug_tmax2006; mtry=7,min.node.size=8; stable VI: NO! horrible stability. 
- rdnbr cut off: 875; variables inlcuded: removed precip_2006 and aug_max_vpd2006 bc of VSURF results; mtry=6,min.node.size=5

- rdnbr cut off: 875; variables inlcuded:tmax, vpd_aug, rain, sr, slope, aspect_transformed, TPI_five, TWI, evi_2006, tslf, cwd, tmin_jan2007, ndmi_2006; mtry=4,min.node.size=3; stable VI: NO! horrible stability.

4 classes
mtry=12,min.node.size=5; stable VI!!; oob not great, TC has lots of confusion with moderate class

3 classes
mtry=9,min.node.size=3; VI is mostly stable!

###... Random Forest
```{r}
for (i in 1:10) {

rfm <- randomForest(rdnbr_qualitative ~., 
                    data = mcf_model3, 
                    #sampsize = c('Not_TC' = 84, 'TC' = 84), 
                    sampsize = c("NLM" = 84,  "High" = 66, "Type Converted" = 84),
                    ntree=1500, 
                    importance=T,
                    replace = T, 
                    mtry = 9,
                    min.node.size = 3)
# OOB and importance
rfm
#importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

}

```

###... ROC/AUC
```{r}
# code: https://stats.stackexchange.com/questions/188616/how-can-we-calculate-roc-auc-for-classification-algorithm-such-as-random-forest
rfm_roc <-  roc(mcf_model$rdnbr_tc, rfm$votes[,2])
plot(rfm_roc)
auc(rfm_roc)
```

###... ALE
```{r}
mcf_ale <- mcf_model3[-which(names(mcf_model3) == "rdnbr_qualitative")]

# predictors 
nlm_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf$rdnbr_qualitative, type = "prob", class="NLM")
high_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf$rdnbr_qualitative, type = "prob", class="High")
tc_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf$rdnbr_qualitative, type = "prob", class="Type Converted")

d = 7

#ndmi
nlm1 <- plot(FeatureEffect$new(nlm_predictor, feature = "ndmi_2006")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high1 <- plot(FeatureEffect$new(high_predictor, feature = "ndmi_2006")) + ggtitle("High") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc1 <- plot(FeatureEffect$new(tc_predictor, feature = "ndmi_2006")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm1, high1, tc1, nrow=1)
ggsave(here::here("Graph Images", "mcf ndmi.jpg"))

#solar radiation
nlm2 <- plot(FeatureEffect$new(nlm_predictor, feature = "sr")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high2 <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("High") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc2 <- plot(FeatureEffect$new(tc_predictor, feature = "sr")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm2, high2, tc2, nrow=1)
ggsave(here::here("Graph Images", "mcf sr.jpg"))

#precip
nlm3 <- plot(FeatureEffect$new(nlm_predictor, feature = "rain")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high3 <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("High") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc3 <- plot(FeatureEffect$new(tc_predictor, feature = "rain")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm3, high3, tc3, nrow=1)
ggsave(here::here("Graph Images", "mcf precip.jpg"))

#tmax

#tmin jan 2007
nlm5 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmin_jan2007")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high5 <- plot(FeatureEffect$new(high_predictor, feature = "tmin_jan2007")) + ggtitle("High") + labs(x="Min 2007 January Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc5 <- plot(FeatureEffect$new(tc_predictor, feature = "tmin_jan2007")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm5, high5, tc5, nrow=1)
ggsave(here::here("Graph Images", "mcf tmin_jan2007.jpg"))

#vpd aug
nlm6 <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high6 <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("High") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc6 <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_aug")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm6, high6, tc6, nrow=1)
ggsave(here::here("Graph Images", "mcf vpd_aug.jpg"))

#cwd
nlm7 <- plot(FeatureEffect$new(nlm_predictor, feature = "cwd")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high7 <- plot(FeatureEffect$new(high_predictor, feature = "cwd")) + ggtitle("High") + labs(x="CWD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc7 <- plot(FeatureEffect$new(tc_predictor, feature = "cwd")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm7, high7, tc7, nrow=1)
ggsave(here::here("Graph Images", "mcf cwd.jpg"))

#evi
nlm8 <- plot(FeatureEffect$new(nlm_predictor, feature = "evi_2006")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high8 <- plot(FeatureEffect$new(high_predictor, feature = "evi_2006")) + ggtitle("High") + labs(x="2006 EVI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc8 <- plot(FeatureEffect$new(tc_predictor, feature = "evi_2006")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm8, high8, tc8, nrow=1)
ggsave(here::here("Graph Images", "mcf evi.jpg"))

#slope
nlm9 <- plot(FeatureEffect$new(nlm_predictor, feature = "slope")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high9 <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("High") + labs(x="Slope")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc9 <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm9, high9, tc9, nrow=1)
ggsave(here::here("Graph Images", "mcf slope.jpg"))

#tslf
nlm10 <- plot(FeatureEffect$new(nlm_predictor, feature = "tslf")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high10 <- plot(FeatureEffect$new(high_predictor, feature = "tslf")) + ggtitle("High") + labs(x="TSLF")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc10 <- plot(FeatureEffect$new(tc_predictor, feature = "tslf")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm10, high10, tc10, nrow=1)
ggsave(here::here("Graph Images", "mcf tslf.jpg"))

#twi
nlm11 <- plot(FeatureEffect$new(nlm_predictor, feature = "TWI")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high11 <- plot(FeatureEffect$new(high_predictor, feature = "TWI")) + ggtitle("High") + labs(x="TWI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc11 <- plot(FeatureEffect$new(tc_predictor, feature = "TWI")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm11, high11, tc11, nrow=1)
ggsave(here::here("Graph Images", "mcf twi.jpg"))

#aspect
nlm12 <- plot(FeatureEffect$new(nlm_predictor, feature = "aspect_transformed")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))  + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high12 <- plot(FeatureEffect$new(high_predictor, feature = "aspect_transformed")) + ggtitle("High") + labs(x="Aspect")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))  + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc12 <- plot(FeatureEffect$new(tc_predictor, feature = "aspect_transformed")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))  + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm12, high12, tc12, nrow=1)
ggsave(here::here("Graph Images", "mcf Aspect.jpg"))

ggarrange(nlm1, high1, tc1, tc1,nlm2, high2, tc2,tc1,nlm3, high3, tc3,tc1, nrow =3, ncol =4)
ggsave(here::here("Graph Images", "mcf ndm_sr_precip6.jpg"))

#ggarrange(nlm5, high5, tc5, nlm6, high6, tc6, nlm7, high7, tc7)

# plot all variables in the model (for just one mortality class): had to specify all the individual variables in the model to plot from the train data frame
# all_rfm_variables <- FeatureEffects$new(high_predictor, method = "ale", features = c('tmax', 'vpd_aug', 'cwd', 'rain', 'insolation', 'slope', 'aspect_transformed', 'TWI', 'ndmi_2006', 'evi_2006', 'years_since_burn', 'tmin_jan2007'))
# all_rfm_variables$plot()
# ggsave("ale4.jpg", height = 9, width = 9)

```

###... ALE
```{r}

mcf_ale <- mcf_model3[-which(names(mcf_model3) == "rdnbr_qualitative")]

# predictors 
nlm_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_qualitative, type = "prob", class="ULM")
high_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_qualitative, type = "prob", class="High")
tc_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_qualitative, type = "prob", class="Type Converted")

d = 7
#ndmi
nlm1 <- plot(FeatureEffect$new(nlm_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high1 <- plot(FeatureEffect$new(high_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc1 <- plot(FeatureEffect$new(tc_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm1, high1, tc1, nrow=1)
##ggsave(here::here("Graph Images", "mcf ndmi.jpg"))
#theme (panel.background = element_rect(width = 5))

#solar radiation
nlm2 <- plot(FeatureEffect$new(nlm_predictor, feature = "sr")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high2 <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc2 <- plot(FeatureEffect$new(tc_predictor, feature = "sr")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
##ggsave(here::here("Graph Images", "mcf sr.jpg"))

#precip
nlm3 <- plot(FeatureEffect$new(nlm_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high3 <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc3 <- plot(FeatureEffect$new(tc_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
##ggsave(here::here("Graph Images", "mcf precip.jpg"))

#evi
nlm4 <- plot(FeatureEffect$new(nlm_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high4 <- plot(FeatureEffect$new(high_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc4 <- plot(FeatureEffect$new(tc_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf evi.jpg"))

#tmax
nlm5 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high5 <- plot(FeatureEffect$new(high_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc5 <- plot(FeatureEffect$new(tc_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf tmax.jpg"))

#tmin jan 2007
nlm6 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high6 <- plot(FeatureEffect$new(high_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc6 <- plot(FeatureEffect$new(tc_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf tmin_jan2007.jpg"))

#vpd aug
nlm7 <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
high7 <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc7 <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm7, high7, tc7, nrow=1)
#ggsave(here::here("Graph Images", "mcf vpd_aug.jpg"))

#cwd
nlm8 <- plot(FeatureEffect$new(nlm_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high8 <- plot(FeatureEffect$new(high_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc8 <- plot(FeatureEffect$new(tc_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf cwd.jpg"))

#slope
nlm9 <- plot(FeatureEffect$new(nlm_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high9 <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc9 <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf slope.jpg"))

#tslf
nlm10 <- plot(FeatureEffect$new(nlm_predictor, feature = "tslf")) + ggtitle("") + labs(x="TSLF")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high10 <- plot(FeatureEffect$new(high_predictor, feature = "tslf")) + ggtitle("") + labs(x="TSLF")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc10 <- plot(FeatureEffect$new(tc_predictor, feature = "tslf")) + ggtitle("") + labs(x="TSLF")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf tslf.jpg"))

#twi
nlm11 <- plot(FeatureEffect$new(nlm_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high11 <- plot(FeatureEffect$new(high_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc11 <- plot(FeatureEffect$new(tc_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf twi.jpg"))

#aspect
nlm12 <- plot(FeatureEffect$new(nlm_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high12 <- plot(FeatureEffect$new(high_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc12 <- plot(FeatureEffect$new(tc_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf Aspect.jpg"))

ggarrange(nlm1, nlm2,nlm3,nlm4,nlm5,nlm6,nlm7,nlm8,nlm9,nlm10,nlm11,nlm12, nrow = 3, ncol=4)
ggsave(here::here("Graph Images", "mcf all ulm_no sampsize.jpg"), width=8, height = 8)

ggarrange(high1,high2,high3,high4,high5,high6,high7,high8,high9,high10,high11,high12, nrow=3, ncol=4)
#ggsave(here::here("Graph Images", "mcf all high_no sampsize.jpg"), width=8, height = 8)

ggarrange(tc1,tc2,tc3,tc4,tc5,tc6,tc7,tc8,tc9,tc10,tc11,tc12, nrow=3, ncol=4)
ggsave(here::here("Graph Images", "mcf all tc_no sampsize.jpg"), width=8, height = 8)

# plot all variables in the model (for just one mortality class): had to specify all the individual variables in the model to plot from the train data frame
# all_rfm_variables <- FeatureEffects$new(tc_predictor, method = "ale", features = c('tmax', 'vpd_aug', 'cwd', 'rain', 'sr', 'slope', 'aspect_transformed', 'TWI', 'ndmi_2006', 'evi_2006', 'tslf', 'tmin_jan2007'))
# all_rfm_variables$plot()
# ggsave("mcf tc.jpg", height = 9, width = 9)

```

###... ALE
```{r}

mcf_ale <- mcf_model[-which(names(mcf_model) == "rdnbr_tc")]

# predictors 
noTC_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_tc, type = "prob", class="Not_TC")
TC_predictor <- Predictor$new(rfm, data = mcf_ale, y = mcf_ale$rdnbr_tc, type = "prob", class="TC")

#ndvi
Low <- plot(FeatureEffect$new(noTC_predictor, feature = "vpd_aug")) + ggtitle("Not Type Converted") + labs(x="NDVI")
high <- plot(FeatureEffect$new(TC_predictor, feature = "vpd_aug")) + ggtitle("Type Converted") + labs(x="NDVI")
ggarrange(Low, high)

#temp
Low <- plot(FeatureEffect$new(noTC_predictor, feature = "rain")) + ggtitle("Not Type Converted")+ labs(x="Precipitation")
high <- plot(FeatureEffect$new(TC_predictor, feature = "rain")) + ggtitle("Type Converted") + labs(x="Precipitation")
ggarrange(Low, high)

#precip
Low <- plot(FeatureEffect$new(noTC_predictor, feature = "evi_2006")) + ggtitle("Not Type Converted") + labs(x="Vegetation Outside")
high <- plot(FeatureEffect$new(TC_predictor, feature = "evi_2006")) + ggtitle("Type Converted")+ labs(x="Vegetation Outside")
ggarrange(Low, high)

#precip
Low <- plot(FeatureEffect$new(noTC_predictor, feature = "ndmi_2006")) + ggtitle("Not Type Converted") + labs(x="Vegetation Inside")
high <- plot(FeatureEffect$new(TC_predictor, feature = "ndmi_2006")) + ggtitle("Type Converted") + labs(x="Vegetation Inside")
ggarrange(Low, high)

# plot all variables in the model (for just one mortality class): had to specify all the individual variables in the model to plot from the train data frame
# all_rfm_variables <- FeatureEffects$new(high_predictor, method = "ale", features = c('tmax', 'vpd_aug', 'cwd', 'rain', 'insolation', 'slope', 'aspect_transformed', 'TWI', 'ndmi_2006', 'evi_2006', 'years_since_burn', 'tmin_jan2007'))
# all_rfm_variables$plot()
# ggsave("ale4.jpg", height = 9, width = 9)
```
interesting graphs for evi and ndmi: high evi (which represents denser veg) is associated with low mortality, which is counterintuative, but higher ndmi (which represents moist veg??) is also associated with low mortality.

## Coulter Pines

#### Load data

```{r}
pico <- read_csv("coulter_env_modified.csv") %>% 
  mutate(rdnbr_qualitative = case_when(
    rdnbr_2007 <= 640 ~ "NLM",
    rdnbr_2007 < 850 ~ "High",
    rdnbr_2007 >= 850 ~ "Type Converted")) %>%
  mutate(rdnbr_qualitative = as.factor(rdnbr_qualitative),
         rdnbr_qualitative = fct_relevel(rdnbr_qualitative, levels=c("NLM","High", "Type Converted"))) %>% 
  #filter(Mortality != "bigcone") %>% # 60 stands actually have big cones not pico
  dplyr::select(-Mortality, -type_converted, -ndvi_2006) %>% 
  na.omit(pico) %>%  # 56 stands have NA values (probs all very small stands)
   mutate(rdnbr_qualitative2 = case_when(
    rdnbr_qualitative %in% "High" ~ "High",
    rdnbr_qualitative %in% c("Moderate", "Low", "None") ~ "NLM")) %>% 
  mutate(rdnbr_qualitative2 = as.factor(rdnbr_qualitative2)) %>% 
  mutate(rdnbr_qualitative2= fct_relevel(rdnbr_qualitative2, levels=c("NLM", "High"))) %>% 
  mutate(aspect_transformed = (1-cos((2*pi*aspect_30m)/360))/2) %>% 
  mutate(rdnbr_tc = case_when(
    rdnbr_2007 < 850 ~ "Not_TC",
    rdnbr_2007 >= 850 ~ "TC"
  )) %>% 
  mutate(rdnbr_tc = as.factor(rdnbr_tc)) %>% 
  mutate(rdnbr_tc= fct_relevel(rdnbr_tc, levels=c("Not_TC", "TC"))) %>% 
  rename_all(funs(str_replace(., "_30m", "")))

pico %>% group_by(rdnbr_qualitative) %>% dplyr::count()

pico_tc_samplesize <- pico %>% group_by(rdnbr_tc) %>% dplyr::count() # 
pico_tc_samplesize <- pico %>% group_by(rdnbr_qualitative) %>% dplyr::count() # nlm = 324, high=120, tc=595
pico_tc_samplesize <- pico %>% group_by(rdnbr_qualitative) %>% dplyr::summarise(acres = sum(acres)) %>% dplyr::summarise(total_acres = sum(acres))
mod1 <- lm(dem~tmin_jan2007, data=pico) # R2=0.368
mod2 <- lm(dem~cwd, data=pico) # R2=0.68
mod3 <- lm(dem~rain, data=pico) # R2=0.475
mod4 <- lm(dem~tmax, data=pico) # R2=0.86
mod5 <- lm(dem~vpd_aug, data=pico) # R2=0.455

pico_elev <- pico %>% dplyr::summarise(elev_mean = mean(dem),
                                       min_elev = min(dem),
                                       max_elev =max(dem))

```

###... Stepwise Variable Selection (VSURF)

```{r}

#Extract prediction and response variables
pico.predictors <- pico_model[,2:ncol(pico_model)]
pico.response <- pico_model[,1]

#Run stepwise variable selection algorithm
pico.vsurf <- VSURF(x = pico.predictors, y = pico.response)

#Get names of selected variables
names(pico.predictors[pico.vsurf$varselect.thres])
names(pico.predictors[pico.vsurf$varselect.interp])
names(pico.predictors[pico$varselect.pred])

```
get error for some reason, but not with the MCF df :(

###... Tuning Parameters

source: https://arxiv.org/pdf/1804.03515.pdf (see digital page 9 for tuneRanger)

options for the measure function: https://mlr.mlr-org.com/articles/tutorial/measures.html

```{r}
set.seed(1127)

# select variables to be used in the model
pico_model <- pico %>% dplyr::select(rdnbr_qualitative, tmax, vpd_aug, cwd, rain, insolation, slope, aspect_transformed, TWI, ndmi_2006, evi_2006, years_since_burn, tmin_jan2007, ndmi_2006)

# run tuneRanger
pico_task <- makeClassifTask(data = pico_model2, target = "rdnbr_tc")
tune_output <- tuneRanger(pico_task, 
                          measure = list(mmce), 
                          num.trees = 1500, 
                          tune.parameters = c("mtry", "min.node.size"))

# recommended parameter settings
# get best parameter values
tune_output$model

## other 
# select variables to be used in the model
pico_model3 <- pico %>% dplyr::select(rdnbr_qualitative, tmax, vpd_aug, cwd, rain, insolation, slope, aspect_transformed, TWI, ndmi_2006, evi_2006, years_since_burn, tmin_jan2007)

# run tuneRanger
pico_task <- makeClassifTask(data = pico_model3, target = "rdnbr_qualitative")
tune_output <- tuneRanger(pico_task, 
                          measure = list(mmce), 
                          num.trees = 1500, 
                          tune.parameters = c("mtry", "min.node.size"))

# recommended parameter settings
# get best parameter values
tune_output$model

```

rdnbr = 850: tmax, vpd_aug, cwd, rain, insolation, slope, aspect_transformed, TWI, ndmi_2006, evi_2006, years_since_burn, tmin_jan2007, ndmi_2006; mtry=3,min.node.size=2

3 class: mtry=1,min.node.size=2; VI stable!

###... Random Forest
```{r}

for (i in 1:10) {

rfm <- randomForest(rdnbr_qualitative ~., 
                    data = pico_model3, 
                    #sampsize = c('Not_TC' = 444, 'TC' = 444), 
                    sampsize = c("NLM" = 120,  "High" = 120, "Type Converted" = 120),
                    ntree=1500, 
                    importance=T,
                    replace = T, 
                    mtry = 1,
                    min.node.size = 2)
# OOB and importance
rfm
#importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

}

#ggplot(pico_model, aes(x=tmin_jan, y=tmin_jan2007)) + geom_point()
#ggplot(pico_model, aes(x=tmax, y=aug_tmax2006)) + geom_point()

```

###... ALE
```{r}
pico_ale <- pico_model3[-which(names(pico_model3) == "rdnbr_qualitative")]

# predictors 
nlm_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_model3$rdnbr_qualitative, type = "prob", class="NLM")
high_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_model3$rdnbr_qualitative, type = "prob", class="High")
tc_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_model3$rdnbr_qualitative, type = "prob", class="Type Converted")

d = 7

#ndmi
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "ndmi_2006")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "ndmi_2006")) + ggtitle("High") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "ndmi_2006")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "pico ndmi.jpg"))
#theme (panel.background = element_rect(width = 5))

#solar radiation
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "insolation")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "insolation")) + ggtitle("High") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "insolation")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "pico sr.jpg"))

#precip
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "rain")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("High") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "rain")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "pico precip.jpg"))

#tmax
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "tmax")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "tmax")) + ggtitle("High") + labs(x="Max August Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "tmax")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "pico tmax.jpg"))

#tmin jan 2007
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "tmin_jan2007")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "tmin_jan2007")) + ggtitle("High") + labs(x="Min 2007 January Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "tmin_jan2007")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "pico tmin_jan2007.jpg"))

#vpd aug
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("High") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_aug")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "pico vpd_aug.jpg"))

#cwd
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "cwd")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "cwd")) + ggtitle("High") + labs(x="CWD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "cwd")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "pico cwd.jpg"))

#evi
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "evi_2006")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "evi_2006")) + ggtitle("High") + labs(x="2006 EVI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "evi_2006")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "pico evi.jpg"))

#slope
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "slope")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("High") + labs(x="Slope")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "pico slope.jpg"))

#tslf
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "years_since_burn")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "years_since_burn")) + ggtitle("High") + labs(x="TSLF")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "years_since_burn")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "pico tslf.jpg"))

#twi
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "TWI")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "TWI")) + ggtitle("High") + labs(x="TWI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "TWI")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "pico twi.jpg"))

#aspect
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "aspect_transformed")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "aspect_transformed")) + ggtitle("High") + labs(x="Aspect")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "aspect_transformed")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "pico Aspect.jpg"))

# plot all variables in the model (for just one mortality class): had to specify all the individual variables in the model to plot from the train data frame
# all_rfm_variables <- FeatureEffects$new(high_predictor, method = "ale", features = c('tmax', 'vpd_aug', 'cwd', 'rain', 'insolation', 'slope', 'aspect_transformed', 'TWI', 'ndmi_2006', 'evi_2006', 'years_since_burn', 'tmin_jan2007'))
# all_rfm_variables$plot()
# ggsave("ale4.jpg", height = 9, width = 9)

```

###... ALE
```{r}

pico_ale <- pico_model3[-which(names(pico_model3) == "rdnbr_qualitative")]

# predictors 
nlm_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_qualitative, type = "prob", class="NLM")
high_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_qualitative, type = "prob", class="High")
tc_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_qualitative, type = "prob", class="Type Converted")

d = 7
#ndmi
nlm1 <- plot(FeatureEffect$new(nlm_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
high1 <- plot(FeatureEffect$new(high_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc1 <- plot(FeatureEffect$new(tc_predictor, feature = "ndmi_2006")) + ggtitle("") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
##ggsave(here::here("Graph Images", "mcf ndmi.jpg"))
#theme (panel.background = element_rect(width = 5))

#solar radiation
nlm2 <- plot(FeatureEffect$new(nlm_predictor, feature = "insolation")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
high2 <- plot(FeatureEffect$new(high_predictor, feature = "insolation")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))#+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc2 <- plot(FeatureEffect$new(tc_predictor, feature = "insolation")) + ggtitle("") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.15, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))#+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
##ggsave(here::here("Graph Images", "mcf sr.jpg"))

#precip
nlm3 <- plot(FeatureEffect$new(nlm_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
high3 <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2))#+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc3 <- plot(FeatureEffect$new(tc_predictor, feature = "rain")) + ggtitle("") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.2,0.2))#+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
##ggsave(here::here("Graph Images", "mcf precip.jpg"))

#evi
nlm4 <- plot(FeatureEffect$new(nlm_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
high4 <- plot(FeatureEffect$new(high_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc4 <- plot(FeatureEffect$new(tc_predictor, feature = "evi_2006")) + ggtitle("") + labs(x="EVI 2006")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf evi.jpg"))

#tmax
nlm5 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
high5 <- plot(FeatureEffect$new(high_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc5 <- plot(FeatureEffect$new(tc_predictor, feature = "tmax")) + ggtitle("") + labs(x="Max August Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf tmax.jpg"))

#tmin jan 2007
nlm6 <- plot(FeatureEffect$new(nlm_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
high6 <- plot(FeatureEffect$new(high_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc6 <- plot(FeatureEffect$new(tc_predictor, feature = "tmin_jan2007")) + ggtitle("") + labs(x="Min 2007 Jan Temp")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf tmin_jan2007.jpg"))

#vpd aug
nlm7 <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
high7 <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc7 <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_aug")) + ggtitle("") + labs(x="Max August VPD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf vpd_aug.jpg"))

#cwd
nlm8 <- plot(FeatureEffect$new(nlm_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
high8 <- plot(FeatureEffect$new(high_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc8 <- plot(FeatureEffect$new(tc_predictor, feature = "cwd")) + ggtitle("") + labs(x="CWD")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf cwd.jpg"))

#slope
nlm9 <- plot(FeatureEffect$new(nlm_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
high9 <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc9 <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("") + labs(x="Slope")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf slope.jpg"))

#tslf
nlm10 <- plot(FeatureEffect$new(nlm_predictor, feature = "years_since_burn")) + ggtitle("") + labs(x="TSLF")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
high10 <- plot(FeatureEffect$new(high_predictor, feature = "years_since_burn")) + ggtitle("") + labs(x="TSLF")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc10 <- plot(FeatureEffect$new(tc_predictor, feature = "years_since_burn")) + ggtitle("") + labs(x="TSLF")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf tslf.jpg"))

#twi
nlm11 <- plot(FeatureEffect$new(nlm_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
high11 <- plot(FeatureEffect$new(high_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc11 <- plot(FeatureEffect$new(tc_predictor, feature = "TWI")) + ggtitle("") + labs(x="TWI")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf twi.jpg"))

#aspect
nlm12 <- plot(FeatureEffect$new(nlm_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
high12 <- plot(FeatureEffect$new(high_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")+ theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc12 <- plot(FeatureEffect$new(tc_predictor, feature = "aspect_transformed")) + ggtitle("") + labs(x="Aspect")  + theme_classic() + scale_y_continuous(lim = c(-0.15,0.2)) #+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
#ggsave(here::here("Graph Images", "mcf Aspect.jpg"))

ggarrange(nlm1, nlm2,nlm3,nlm4,nlm5,nlm6,nlm7,nlm8,nlm9,nlm10,nlm11,nlm12, nrow = 3, ncol=4)
ggsave(here::here("Graph Images", "pico all ulm.jpg"), width=8, height = 8)

ggarrange(high1,high2,high3,high4,high5,high6,high7,high8,high9,high10,high11,high12, nrow=3, ncol=4)
ggsave(here::here("Graph Images", "pico all high.jpg"), width=8, height = 8)

ggarrange(tc1,tc2,tc3,tc4,tc5,tc6,tc7,tc8,tc9,tc10,tc11,tc12, nrow=3, ncol=4)
ggsave(here::here("Graph Images", "pico all tc.jpg"), width=8, height = 8)

# plot all variables in the model (for just one mortality class): had to specify all the individual variables in the model to plot from the train data frame
# all_rfm_variables <- FeatureEffects$new(tc_predictor, method = "ale", features = c('tmax', 'vpd_aug', 'cwd', 'rain', 'sr', 'slope', 'aspect_transformed', 'TWI', 'ndmi_2006', 'evi_2006', 'tslf', 'tmin_jan2007'))
# all_rfm_variables$plot()
# ggsave("mcf tc.jpg", height = 9, width = 9)

```

###... ALE
```{r}

pico_ale <- pico_model2[-which(names(pico_model2) == "rdnbr_tc")]

# predictors 
noTC_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_tc, type = "prob", class="Not_TC")
TC_predictor <- Predictor$new(rfm, data = pico_ale, y = pico_ale$rdnbr_tc, type = "prob", class="TC")

#ndvi
Low <- plot(FeatureEffect$new(noTC_predictor, feature = "vpd_aug")) + ggtitle("Not Type Converted") + labs(x="NDVI")
high <- plot(FeatureEffect$new(TC_predictor, feature = "vpd_aug")) + ggtitle("Type Converted") + labs(x="NDVI")
ggarrange(Low, high)

#temp
Low <- plot(FeatureEffect$new(noTC_predictor, feature = "rain")) + ggtitle("Not Type Converted")+ labs(x="Precipitation")
high <- plot(FeatureEffect$new(TC_predictor, feature = "rain")) + ggtitle("Type Converted") + labs(x="Precipitation")
ggarrange(Low, high)

#precip
Low <- plot(FeatureEffect$new(noTC_predictor, feature = "evi_2006")) + ggtitle("Not Type Converted") + labs(x="Vegetation Outside")
high <- plot(FeatureEffect$new(TC_predictor, feature = "evi_2006")) + ggtitle("Type Converted")+ labs(x="Vegetation Outside")
ggarrange(Low, high)

#precip
Low <- plot(FeatureEffect$new(noTC_predictor, feature = "ndmi_2006")) + ggtitle("Not Type Converted") + labs(x="Vegetation Inside")
high <- plot(FeatureEffect$new(TC_predictor, feature = "ndmi_2006")) + ggtitle("Type Converted") + labs(x="Vegetation Inside")
ggarrange(Low, high)

# plot all variables in the model (for just one mortality class): had to specify all the individual variables in the model to plot from the train data frame
all_rfm_variables <- FeatureEffects$new(noTC_predictor, method = "ale", features = c('tmax', 'vpd_aug', 'cwd', 'rain', 'insolation', 'slope', 'aspect_transformed', 'TWI', 'ndmi_2006', 'evi_2006', 'years_since_burn', 'tmin_jan2007'))
all_rfm_variables$plot()
ggsave("pico_no_tc2.jpg", height = 9, width = 9)
```

#COMBINED COulter and MCF data

##... prep data
```{r}
# rename some variables to match with MCF
pico_mod <- pico_model3 %>% rename(sr = insolation, tslf = years_since_burn)

# combine mcf and pico

all_data <- rbind(pico_mod, mcf_model3)

# samp size
all_data %>%  group_by(rdnbr_qualitative) %>% dplyr::count()

```

###... Tuning Parameters

source: https://arxiv.org/pdf/1804.03515.pdf (see digital page 9 for tuneRanger)

options for the measure function: https://mlr.mlr-org.com/articles/tutorial/measures.html

```{r}
set.seed(1127)

# run tuneRanger
all_task <- makeClassifTask(data = all_data, target = "rdnbr_qualitative")
tune_output <- tuneRanger(all_task, 
                          measure = list(mmce), 
                          num.trees = 1500, 
                          tune.parameters = c("mtry", "min.node.size"))

# recommended parameter settings
tune_output$model

```
mtry=3,min.node.size=3

###... Random Forest
```{r}

for (i in 1:10) {

rfm <- randomForest(rdnbr_qualitative ~., 
                    data = all_data, 
                    sampsize = c("NLM" = 186,  "High" = 186, "Type Converted" = 186),
                    ntree=1500, 
                    importance=T,
                    replace = T, 
                    mtry = 3,
                    min.node.size = 3)
# OOB and importance
rfm
#importance(rfm)
#plot to visualize variable importance
varImpPlot(rfm)

}

```

###... ALE
```{r}
all_ale <- all_data[-which(names(all_data) == "rdnbr_qualitative")]

# predictors 
nlm_predictor <- Predictor$new(rfm, data = all_ale, y = all_data$rdnbr_qualitative, type = "prob", class="NLM")
high_predictor <- Predictor$new(rfm, data = all_ale, y = all_data$rdnbr_qualitative, type = "prob", class="High")
tc_predictor <- Predictor$new(rfm, data = all_ale, y = all_data$rdnbr_qualitative, type = "prob", class="Type Converted")

d = 7

#ndmi
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "ndmi_2006")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.19, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "ndmi_2006")) + ggtitle("High") + labs(x="NDMI 2006") + theme_classic() + scale_y_continuous(lim = c(-0.19, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "ndmi_2006")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.19, 0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "all ndmi.jpg"))
#theme (panel.background = element_rect(width = 5))

#solar radiation
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "sr")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.19, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "sr")) + ggtitle("High") + labs(x="Solar Radiation") + theme_classic() + scale_y_continuous(lim = c(-0.19, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "sr")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.19, 0.2)) +  theme(axis.text.x = element_text(angle = 15, hjust = 0.85))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "all sr.jpg"))

#precip
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "rain")) + ggtitle("ULM") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "rain")) + ggtitle("High") + labs(x="Annual Precipitation") + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "rain")) + ggtitle("Type Converted") + labs(x="") + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2))+ theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "all precip.jpg"))

#tmax
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "tmax")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.21)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "tmax")) + ggtitle("High") + labs(x="Max August Temp") + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "tmax")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "all tmax.jpg"))

#tmin jan 2007
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "tmin_jan2007")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "tmin_jan2007")) + ggtitle("High") + labs(x="Min 2007 January Temp")+ theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "tmin_jan2007")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "all tmin_jan2007.jpg"))

#vpd aug
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "vpd_aug")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "vpd_aug")) + ggtitle("High") + labs(x="Max August VPD")+ theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "vpd_aug")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "all vpd_aug.jpg"))

#cwd
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "cwd")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "cwd")) + ggtitle("High") + labs(x="CWD")+ theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "cwd")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "all cwd.jpg"))

#evi
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "evi_2006")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "evi_2006")) + ggtitle("High") + labs(x="2006 EVI")+ theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "evi_2006")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "all evi.jpg"))

#slope
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "slope")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "slope")) + ggtitle("High") + labs(x="Slope")+ theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "slope")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "all slope.jpg"))

#tslf
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "tslf")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "tslf")) + ggtitle("High") + labs(x="TSLF")+ theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "tslf")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "all tslf.jpg"))

#twi
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "TWI")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "TWI")) + ggtitle("High") + labs(x="TWI")+ theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "TWI")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "all twi.jpg"))

#aspect
nlm <- plot(FeatureEffect$new(nlm_predictor, feature = "aspect_transformed")) + ggtitle("ULM") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
high <- plot(FeatureEffect$new(high_predictor, feature = "aspect_transformed")) + ggtitle("High") + labs(x="Aspect")+ theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
tc <- plot(FeatureEffect$new(tc_predictor, feature = "aspect_transformed")) + ggtitle("Type Converted") + labs(x="")  + theme_classic() + scale_y_continuous(lim = c(-0.19,0.2)) + theme(plot.margin = unit(c(d,1,1,1), "lines"))
ggarrange(nlm, high, tc, nrow=1)
ggsave(here::here("Graph Images", "all Aspect.jpg"))

# plot all variables in the model (for just one mortality class): had to specify all the individual variables in the model to plot from the train data frame
# all_rfm_variables <- FeatureEffects$new(high_predictor, method = "ale", features = c('tmax', 'vpd_aug', 'cwd', 'rain', 'insolation', 'slope', 'aspect_transformed', 'TWI', 'ndmi_2006', 'evi_2006', 'years_since_burn', 'tmin_jan2007'))
# all_rfm_variables$plot()
# ggsave("ale4.jpg", height = 9, width = 9)

```